Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2021-09-14 vertical-thermal-review.Rmd
isoprenes

**The vertical structuring of leaf isoprene emissions may indicate the importance of temporal variability in thermal conditions as distinct from the long-term average.** Isoprene gas emission from tree leaves is associated with photosynthetic thermal tolerance [@taylorCapacityEmitIsoprene2019; @sharkeyIsopreneEmissionPlants2008], and within species, emission rates tend to increase toward brighter and hotter microenvironments [@niinemetsEmissionFactorVolatile2010]. Across landscapes, emitting species increase in relative abundance toward hotter climates, exceeding half of trees in warm tropical forests [@taylorIsopreneEmissionStructures2018a]. In contrast, a recent inventory in an Amazon forest showed that both the abundance of emitting species and average emission rates were highest in the mid-canopy region instead of the hotter and brighter upper canopy [@taylorNewFieldInstrument2021]. This observation is consistent with the hypothesis that temperature sensitive isoprene emissions enable real-time acclimation to rapidly changing leaf thermal environments typical of the mid-canopy region [@sharkeyIsopreneEmissionPlants2008], due to the sun's passage across a dynamically structured canopy surface. This review focuses primarily on variation in mean conditions across the vertical profile, but future work should seek to understand how temporal dynamics of leaf heating [see @leighInfluenceLeafSize2017] vary through the canopy and influence leaf function.

>>>>

Isoprene gas emission from tree leaves is associated with photosynthetic thermal tolerance [@taylorCapacityEmitIsoprene2019; @sharkeyIsopreneEmissionPlants2008], and within species, emission rates tend to increase toward brighter and hotter microenvironments [@niinemetsEmissionFactorVolatile2010]. 
Across landscapes, emitting species increase in relative abundance toward hotter climates, exceeding half of trees in warm tropical forests [@taylorIsopreneEmissionStructures2018a].
In contrast, vertical structuring of leaf isoprene emissions within forests may indicate the importance of temporal variability in thermal conditions as distinct from the long-term average.
A recent inventory in an Amazon forest showed that both the abundance of emitting species and average emission rates were highest in the mid-canopy region instead of the hotter and brighter upper canopy [@taylorNewFieldInstrument2021]. 
This observation is consistent with the hypothesis that temperature sensitive isoprene emissions enable real-time acclimation to rapidly changing leaf thermal environments typical of the mid-canopy region [@sharkeyIsopreneEmissionPlants2008], due to the sun's passage across a dynamically structured canopy surface. 
This review focuses primarily on variation in mean conditions across the vertical profile, but future work should seek to understand how temporal dynamics of leaf heating [see @leighInfluenceLeafSize2017] vary through the canopy and influence leaf function.
